Individual heterogeneity in functional topography of the default-mode network in medial prefrontal cortex

### 2019-03-29 Manuscript.Rmd
update to bibliography and tweaks to manuscript before knitting

---
title: 'Individual heterogeneity in functional topography of the default-mode network in medial prefrontal cortex'
author: "Claudio Toro-Serey, Sean Tobyne, & Joseph T. McGuire"
output:
  pdf_document: 
    keep_tex: true
  word_document: 
    reference_docx: Style_reference_draft.docx
  html_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

  Human medial prefrontal cortex (mPFC) and posterior cingulate cortex (PCC) are jointly associated with a large set of disparate cognitive functions [@Hiser2018; @Kragel2018]. One such function is subjective valuation during economic decision making [@Bartra2013; @Clithero2014; @Hiser2018; @Kable2007; @Levy2011]. Both mPFC and PCC (together with ventral striatum) consistently show greater BOLD activity in response to more highly valued choice prospects and outcomes, relative to prospects and outcomes that are less highly valued. These regions are also implicated in a set of functions associated with the default-mode network (DMN). DMN effects include a decrease in BOLD activity during cognitively demanding tasks compared with less-demanding task conditions or periods of rest [@McKiernan2003; @Laird2009], as well as a distinctive pattern of inter-region correlations in resting-state fMRI data [@Fox2005; @Yeo2011].

  Coordinate-based meta-analyses of the fMRI literature have demonstrated that valuation and DMN effects occur in overlapping and largely indistinguishable regions of ventral mPFC [@Acikalin2017]. The two sets of effects diverge in other brain areas; a region of the striatum is preferentially associated with valuation, whereas bilateral temporoparietal junction is linked with the DMN. Within PCC and adjacent precuneus, a more posterior subregion appears DMN-specific, whereas a more anterior subregion is implicated in both effects. One possible interpretation of these findings is that mPFC and anterior PCC operate as common nodes within the distributed brain systems that subserve valuation and DMN effects. The potential existence of common nodes has primpted theorizing about posible cognitive-level overlap between valuation and DMN-related processes. For example, the two sets of processes might share elements of self-referential cognition, episodic memory, mental simulation, or monitoring and regulation of internal states [@Northoff2012; @Clithero2014; @Acikalin2017; Reddan2018].

  However, to draw firm conclusions about functional overlap will require a better understanding of individual-level heterogeneity in topographic patterns of brain activity. A key limitation of group averaging and meta-analysis is that individual-specific functional subdivisions can be blurred and overlooked [@Michalka2015; @Fedorenko2012; Woo2014], potentially exaggerating the degree of overlap across domains. This concern is especially pronounced in mPFC, which is subject to considerable idiosyncratic cortical folding [@Zilles2013; @Mackey2014] and inter-subject functional variability [@Mueller2013a]. An alternative approach is to focus on analyses at the individual-participant level. Individual-level analyses of fMRI data have identified idiosyncratic, reliable, and valid patterns of functional organization that would be blurred in aggregative estimates [@Laumann2015; @Gordon2017; @Gratton2018a; @Tobyne2018]. Subject-specific network arrangements have been found to predict behavioral characteristics [@Kong2018]. Recent work has uncovered fine-grained subdivisions within the DMN through tailored analyses of seed-based functional connectivity [@Braga2017]. It is therefore possible that the diffuse overlap of DMN and valuation effects can be attributed to a lack of spatial resolution, and that the organization of mPFC and PCC would be better understood at the individual level. An important first step, and the goal of the present paper, is to quantify the degree of topographic heterogeneity of DMN effects within mPFC and PCC.

@@

For all analyses, we defined our search space based on the 17-network parcellation proposed by Yeo et al. [-@Yeo2011]. First, we selected vertices on the cortical surface that were contained by the DMN and limbic networks along the medial wall. Next, we overlayed those networks on a parcellated atlas of the human cortical surface (360 regions) [@Glasser2016], and retained a set of parcels that covered approximately the same brain regions. This resulted in a search space that consisted of 40 parcels across hemispheres (Table 1). The search space was naturally divided into two spatially discontiguous clusters that covered PCC and mPFC, facilitating the study of each general region separately.

@@

We used a novel approach to cortical surface parcel-based meta-analysis to assess whether individual parcels within the search space were preferentially associated with valuation or DMN effects. For subjective valuation, we gathered peak brain coordinates of activity from 200 studies that were associated with positive effects for current or prospective rewards [@Bartra2013]. For DMN, we acquired coordinates from a previous meta-analysis of 80 studies that were related to task-deactivation [@Laird2009]. These coordinates represent the areas that exceeded the statistical significance threshold in each original study. For each study, we created an indicator map in standard space (MNI152, 1 mm resolution) which contained values of 1 in a 10 mm radius sphere around each reported activation peak, and values of 0 elsewhere [@Wager2009]. The indicator map for each study was then projected to a standard cortical mesh (fs_average, 32,000 vertices, projfrac-max from 0 to 1 by 0.25) using FreeSurfer's mri_vol2surf [@Fischl2012] (http://surfer.nmr.mgh.harvard.edu/). Any sphere sections that overlapped with the cortex (no lower limit restrictions) were mapped to the corresponding parcel, and we tallied how many studies had positive indicator values in each parcel. Two studies from the subjective valuation corpus were removed because they did not contain cortical voxels, leaving a final number of 198 studies.

@@

All network analyses were performed using the igraph package [@G.2006] in R 3.4.1 [@RTeam2018]. To establish each subject's network, we selected all the vertices contained within the mPFC/PCC search space (n = 4,801 per subject; mPFC = 2854, PCC = 1947) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, in which the nodes were surface vertices and the edges were the correlations among them. Next, we applied Fisher's r to z transformation, and performed a two-sided significance test to identify strong correlations. The resulting p-values were corrected for multiple comparisons (FDR < 0.05), and edges with non-significant correlations were turned to 0 (all other edges retained their respective z-values). **TODO: recompute corrCounts so it also stores the # of pos/neg correlations turned to 0, as well as the number of TRs per subject (add to fMRI deets)**. Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution. Values of 1 of this transformed distribution were returned to 0 in order to avoid the inclusion of these edges in the network analyses. The proportion of remaining edges (i.e. network density) was high for all individuals(mean = `r round(mdAll, digits = 2)`, SD = `r round(sddAll, digits = 3)`). We generated and analyzed network weight matrices at four levels: (1) for each run separately (1200 TRs each); (2) for the concatenated time series for the two runs on each day (2400 TRs total); (3) for each subject's full dataset (4800 TRs total); and (4) on each step of a sliding window analysis (see Partition Evaluation for more details).

@@

We evaluated the performance of the SP algorithm in comparison to a simpler partitioning approach using seed-based functional connectivity. First, we estimated each subject's DMN partition in mPFC based on its vertex-wise functional correlation (Pearson) with the average activity across all vertices in PCC for each day separately. We compared these seed-based maps with the Fiedler vectors produced by SP by examining three sets of spatial correlations across vertices between days: 1) between seed-based maps; 2) between SP-based maps; and 3) between seed- and SP-based maps. These spatial correlations were meant to determine the test/re-test reliability of each approach, as well as the overall level of agreement between them. The two methods were expected to produce somewhat similar results, but the one displaying greater within-subject agreement across days should be preferred (for a discussion on the stability of functional networks see [@Kong2018; @Gratton2018a]). We therefore compared the mean within-subject inter-day correlation coefficients of each method through a paired permutation analysis (5000 iterations).

@@

``` {r Correlation vs community detection: compute, fig.align="center", echo=FALSE, message = FALSE, cache = TRUE}
## Between halves analysis for correlation and communities (mPFC only)
unthreshComm_PFC_halves_plot <- lapply(seq_along(unthreshComm_PFC_halves_plot), function(x) cbind(unthreshComm_PFC_halves_plot[[x]], Correlation_halves_plot[[x]]))
D1 <- unthreshComm_PFC_halves_plot
D2 <- lapply(unthreshComm_PFC_halves_plot, function(data) transform(data, V1 = FiedlerVec)) # the comparison function can only take 1 column name for this stuff
Data <- c(D1,D2)
subjList_halves_plot <- rep(best$SubjID, each = 2)
allHalfCombs <- comparePartitions(Data = Data, MOI = "V1", nSubjects = 40, Index = "Cor", subjNames = c(subjList_halves_plot, subjList_halves_plot))
```

@@

A considerable amount of meta-analytic work has been dedicated to pinpointing the spatial embedding of behavioral constructs in medial prefrontal cortex (mPFC), revealing both dissociable and overlapping activation across psychological phenomena [@DelaVega2016; @Kragel2018; @Hiser2018]. In particular, the effects elicited by the subjective value (SV) of an option and the default mode network (DMN) have been suggested to be inseparable in this area, an overlap that partially extends to posterior cingulate cortex (PCC) [@Acikalin2017; @Bartra2013; @Clithero2014; @Laird2009]. However, recent studies have shown considerable variability of the brain's functional topography across individuals, emphasizing the importance of capturing its organization at the individual subject level, and pointing to limitations in aggregative estimations [@Braga2017; @Gordon2017; @Gratton2018a; @Kong2018; @Laumann2015; @Tobyne2018]. Therefore, determining the distinct functional mapping of psychological processes (and the potential extent of their overlap) first requires: 1) understanding and quantifying the degree of heterogeneity in the intrinsic functional organization of this area in a large group of individuals; and 2) a method to reliably and precisely capture it in isolated subjects.

>>>>

---
title: 'Individual heterogeneity in functional topography of the default-mode network in medial prefrontal cortex'
author: "Claudio Toro-Serey, Sean Tobyne, & Joseph T. McGuire"
output:
  word_document: 
    reference_docx: Style_reference_draft.docx
  pdf_document: 
    keep_tex: true
  html_document: default
bibliography: mPFC_paper.bib
header-includes:
    - \usepackage{setspace}\doublespacing
---

@@

  Human medial prefrontal cortex (mPFC) and posterior cingulate cortex (PCC) are jointly associated with a large set of disparate cognitive functions [@Hiser2018; @Kragel2018]. One such function is subjective valuation during economic decision making [@Bartra2013; @Clithero2014; @Hiser2018; @Kable2007; @Levy2011]. Both mPFC and PCC (together with ventral striatum) consistently show greater BOLD activity in response to more highly valued choice prospects and outcomes, relative to prospects and outcomes that are less highly valued. These regions are also implicated in a set of functions associated with the default-mode network (DMN). DMN effects include a decrease in BOLD activity during cognitively demanding tasks compared with less-demanding task conditions or periods of rest [@Mckiernan2003; @Laird2009], as well as a distinctive pattern of inter-region correlations in resting-state fMRI data [@Fox2005; @Yeo2011].

  Coordinate-based meta-analyses of the fMRI literature have demonstrated that valuation and DMN effects occur in overlapping and largely indistinguishable regions of ventral mPFC [@Acikalin2017]. The two sets of effects diverge in other brain areas; a region of the striatum is preferentially associated with valuation, whereas bilateral temporoparietal junction is linked with the DMN. Within PCC and adjacent precuneus, a more posterior subregion appears DMN-specific, whereas a more anterior subregion is implicated in both effects. One possible interpretation of these findings is that mPFC and anterior PCC operate as common nodes within the distributed brain systems that subserve valuation and DMN effects. The potential existence of common nodes has primpted theorizing about posible cognitive-level overlap between valuation and DMN-related processes. For example, the two sets of processes might share elements of self-referential cognition, episodic memory, mental simulation, or monitoring and regulation of internal states [@Northoff2012; @Clithero2014; @Acikalin2017; @Reddan2018].

  However, to draw firm conclusions about functional overlap will require a better understanding of individual-level heterogeneity in topographic patterns of brain activity. A key limitation of group averaging and meta-analysis is that individual-specific functional subdivisions can be blurred and overlooked [@Fedorenko2012; @Woo2014; @Michalka2015; @Wang2015], potentially exaggerating the degree of overlap across domains. This concern is especially pronounced in mPFC, which is subject to considerable idiosyncratic cortical folding [@Zilles2013; @Mackey2014] and inter-subject functional variability [@Mueller2013a]. An alternative approach is to focus on analyses at the individual-participant level. Individual-level analyses of fMRI data have identified idiosyncratic, reliable, and valid patterns of functional organization that would be blurred in aggregative estimates [@Laumann2015; @Gordon2017; @Gratton2018b; @Tobyne2018]. Subject-specific network arrangements have been found to predict behavioral characteristics [@Kong2018]. Recent work has uncovered fine-grained subdivisions within the DMN through tailored analyses of seed-based functional connectivity [@Braga2017; @Braga2019]. It is therefore possible that the diffuse overlap of DMN and valuation effects can be attributed to a lack of spatial resolution, and that the organization of mPFC and PCC would be better understood at the individual level. An important first step, and the goal of the present paper, is to quantify the degree of topographic heterogeneity of DMN effects within mPFC and PCC.

@@

For all analyses, we defined our search space based on the 17-network parcellation proposed by Yeo et al. [-@Yeo2011]. First, we selected vertices on the cortical surface that were contained by the DMN and limbic networks along the medial wall. Next, we overlayed those networks on a parcellated atlas of the human cortical surface (360 regions) [@Glasser2016], and retained a set of parcels that covered approximately the same brain regions. This resulted in a search space that consisted of 40 parcels across hemispheres (Supplementary Table 1). The search space was naturally divided into two spatially discontiguous clusters that covered PCC and mPFC, facilitating the study of each general region separately.

@@

We used a novel approach to cortical surface parcel-based meta-analysis to assess whether individual parcels within the search space were preferentially associated with valuation or DMN effects. For subjective valuation, we gathered peak brain coordinates of activity from 200 studies that were associated with positive effects for current or prospective rewards [@Bartra2013]. For DMN, we acquired coordinates from a previous meta-analysis of 80 studies that were related to task-deactivation [@Laird2009]. These coordinates represent the areas that exceeded the statistical significance threshold in each original study. For each study, we created an indicator map in standard space (MNI152, 1 mm resolution) which contained values of 1 in a 10 mm radius sphere around each reported activation peak, and values of 0 elsewhere [@Wager2009]. The indicator map for each study was then projected to a standard cortical mesh (fs_average, 32,000 vertices, projfrac-max from 0 to 1 by 0.25) using FreeSurfer's mri_vol2surf [@Fischl1999; @Dale1999] (http://surfer.nmr.mgh.harvard.edu/). Any sphere sections that overlapped with the cortex (no lower limit restrictions) were mapped to the corresponding parcel, and we tallied how many studies had positive indicator values in each parcel. Two studies from the subjective valuation corpus were removed because they did not contain cortical voxels, leaving a final number of 198 studies.

@@

All network analyses were performed using the igraph package [@G.2006] in R 3.4.1 [@RTeam2018]. To establish each subject's network, we selected all the vertices contained within the mPFC/PCC search space (n = 4,801 per subject; mPFC = 2854, PCC = 1947) and computed the Pearson correlation of the time series for every pair of vertices. This produced a weighted network for each subject, in which the nodes were surface vertices and the edges were the correlations among them. Next, we applied Fisher's r to z transformation, and performed a two-sided significance test to identify strong correlations. The resulting p-values were corrected for multiple comparisons (FDR < 0.05), and edges with non-significant correlations were turned to 0 (all other edges retained their respective z-values). **TODO: recompute corrCounts so it also stores the # of pos/neg correlations turned to 0, as well as the number of TRs per subject (add to fMRI details)**. Next, we took the exponential of these values so that all weights were positive while maintaining the ordinal ranks of the original correlation distribution. Values of 1 of this transformed distribution were returned to 0 in order to avoid the inclusion of these edges in the network analyses. The proportion of remaining edges (i.e. network density) was high for all individuals(mean = `r round(mdAll, digits = 2)`, SD = `r round(sddAll, digits = 3)`). We generated and analyzed network weight matrices at four levels: (1) for each run separately (1200 TRs each); (2) for the concatenated time series for the two runs on each day (2400 TRs total); (3) for each subject's full dataset (4800 TRs total); and (4) on each step of a sliding window analysis (see Partition Evaluation for more details).

@@

We evaluated the performance of the SP algorithm in comparison to a simpler partitioning approach using seed-based functional connectivity. First, we estimated each subject's DMN partition in mPFC based on its vertex-wise functional correlation (Pearson) with the average activity across all vertices in PCC for each day separately. We compared these seed-based maps with the Fiedler vectors produced by SP by examining three sets of spatial correlations across vertices between days: 1) between seed-based maps; 2) between SP-based maps; and 3) between seed- and SP-based maps. These spatial correlations were meant to determine the test/re-test reliability of each approach, as well as the overall level of agreement between them. The two methods were expected to produce somewhat similar results, but the one displaying greater within-subject agreement across days should be preferred (for a discussion on the stability of functional networks see [@Kong2018; @Gratton2018b]). We therefore compared the mean within-subject inter-day correlation coefficients of each method through a paired permutation analysis (5000 iterations).

@@

``` {r Correlation vs community detection: compute, fig.align="center", echo=FALSE, message = FALSE, cache = TRUE}
## Between halves analysis for correlation and communities (mPFC only)
# Spearman to avoid issues with the differences in value ranges
unthreshComm_PFC_halves_plot <- lapply(seq_along(unthreshComm_PFC_halves_plot), function(x) cbind(unthreshComm_PFC_halves_plot[[x]], Correlation_halves_plot[[x]]))
D1 <- unthreshComm_PFC_halves_plot
D2 <- lapply(unthreshComm_PFC_halves_plot, function(data) transform(data, V1 = FiedlerVec)) # the comparison function can only take 1 column name for this stuff
Data <- c(D1,D2)
subjList_halves_plot <- rep(best$SubjID, each = 2)
allHalfCombs <- comparePartitions(Data = Data, MOI = "V1", nSubjects = 40, Index = "Cor", subjNames = c(subjList_halves_plot, subjList_halves_plot))

Foo <- "fighters"
```

@@

A considerable amount of meta-analytic work has been dedicated to pinpointing the spatial embedding of behavioral constructs in medial prefrontal cortex (mPFC), revealing both dissociable and overlapping activation across psychological phenomena [@DelaVega2016; @Kragel2018; @Hiser2018]. In particular, the effects elicited by the subjective value (SV) of an option and the default mode network (DMN) have been suggested to be inseparable in this area, an overlap that partially extends to posterior cingulate cortex (PCC) [@Acikalin2017; @Bartra2013; @Clithero2014; @Laird2009]. However, recent studies have shown considerable variability of the brain's functional topography across individuals, emphasizing the importance of capturing its organization at the individual subject level, and pointing to limitations in aggregative estimations [@Braga2017; @Gordon2017; @Gratton2018b; @Kong2018; @Laumann2015; @Tobyne2018]. Therefore, determining the distinct functional mapping of psychological processes (and the potential extent of their overlap) first requires: 1) understanding and quantifying the degree of heterogeneity in the intrinsic functional organization of this area in a large group of individuals; and 2) a method to reliably and precisely capture it in isolated subjects.
